---
title: "EV Power - Lab 4 Project Report"
format: typst
---

# Example Solution 1

## **Part 0: libraries**

```{r}
---
title: "Electric Vehicles and Renewable Energy in the U.S."
author: "Rabia Ali"
date: "2025-10-29"
format: pdf
---

```{r setup, include=FALSE}
# Load required libraries
# Run this once to install missing libraries
install.packages(c("tidyverse", "sf", "tmap", "tigris", "leaflet"))

library(tidyverse)
library(sf)
library(tmap)
library(tigris)
library(leaflet)

# Cache shapefiles for faster loading
options(tigris_use_cache = TRUE)

```

## **Part 1:** **Defining Research Question**

Chosen Question: Electric vehicles reduce direct emissions, but does the electricity used to charge them actually come from clean sources?
What is the share of electricity from renewable sources by state in 2023?

## **Part 2: Data Preparation and Cleaning**

```{r}
# Load renewable energy CSVs

renew_2021 <- read_csv("renew_use_2021.csv")
renew_2022 <- read_csv("renew_use_2022.csv")
renew_2023 <- read_csv("renew_use_2023.csv")

# Load total energy CSVs

total_2021 <- read_csv("total_energy_use_2021.csv")
total_2022 <- read_csv("total_energy_use_2022.csv")
total_2023 <- read_csv("total_energy_use_2023.csv")

# Load average energy price

energy_price <- read_csv("av_energy_price_2021-2023.csv")

# Load EV registrations

ev_regs <- read_csv("ev-registrations-by_state_2023.csv")
# Function to clean column names

clean_names <- function(df) {
df %>%
rename_with(~str_to_lower(.x)) %>%
rename_with(~str_replace_all(.x, " ", "*")) %>%
rename_with(~str_replace_all(.x, "-", "*"))
}

renew_2021 <- clean_names(renew_2021)
renew_2022 <- clean_names(renew_2022)
renew_2023 <- clean_names(renew_2023)
total_2021 <- clean_names(total_2021)
total_2022 <- clean_names(total_2022)
total_2023 <- clean_names(total_2023)
energy_price <- clean_names(energy_price)
ev_regs <- clean_names(ev_regs)

# Standardize state names

standardize_states <- function(df) {
df %>% mutate(state = str_to_title(state))
}

renew_2021 <- standardize_states(renew_2021)
renew_2022 <- standardize_states(renew_2022)
renew_2023 <- standardize_states(renew_2023)
total_2021 <- standardize_states(total_2021)
total_2022 <- standardize_states(total_2022)
total_2023 <- standardize_states(total_2023)
energy_price <- standardize_states(energy_price)
ev_regs <- standardize_states(ev_regs)

```

## **Part 3: Joining / Pivoting Datasets for Analysis**

```{r}
# Combine yearly renewable and total energy data

renew_all <- bind_rows(
renew_2021 %>% mutate(year = 2021),
renew_2022 %>% mutate(year = 2022),
renew_2023 %>% mutate(year = 2023)
)

total_all <- bind_rows(
total_2021 %>% mutate(year = 2021),
total_2022 %>% mutate(year = 2022),
total_2023 %>% mutate(year = 2023)
)

# Merge datasets

energy_data <- renew_all %>%
left_join(total_all, by = c("state", "year")) %>%
left_join(energy_price, by = c("state", "year")) %>%
left_join(ev_regs %>% mutate(year = 2023), by = "state")

# Create new variables

energy_data <- energy_data %>%
mutate(
renewable_share = renewable_use / total_energy_use,
ev_ratio = ev_registrations / total_energy_use
)

```

## **Part 4: Mapping Visualization**

```{r}
# Check first few rows

head(energy_data)
# Load US states shapefile

states_sf <- states(cb = TRUE)

# Filter for 2023

energy_2023 <- energy_data %>% filter(year == 2023)

# Join shapefile with data

map_data <- states_sf %>%
left_join(energy_2023, by = c("NAME" = "state"))

# Static map using tmap

tm_shape(map_data) +
tm_polygons(
"renewable_share",
palette = "Greens",
title = "Renewable Share",
style = "quantile"
) +
tm_layout(title = "Renewable Energy Share by State (2023)")


```

```
